Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2024-10-07 analysis/paper/article.qmd
article qmd for the journal article
starting the results

title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 900
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Academic careers are frequently sought by archaeology graduate students. Job listing websites often serve as the first place for these students when seeking academic positions. We examined tenure-track job advertisements over the past decade to gain insights into the academic job market for archaeologists. Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time. We studied the text of 449 job ads posted from 2013-2023. Our analysis focuses on shifts in archaeological topics and methods requested in job ads. We investigate whether the burden on applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago? We also examine whether there is an increasing trend in job advertisements highlighting diversity and inclusivity, thereby encouraging a broader range of applicants. Additionally, we assess the influence of socio-political factors on the changing focus of research topics in the field. This research aims to assist current and future archaeology students and graduates in better understanding the job market and the requirements of employers, thereby aiding them in effectively preparing for their applications for positions in archaeology.
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

@@

@@

total_number_of_ads_in_our_sample <- nrow(jobdata) # 550
```

We have `r total_number_of_ads_in_our_sample` job advertisements in our sample

```{r}
#| label: fig-how-many-jobs-per-year
#| fig-cap: "Number of tenure-track archaeology faculty job ads posted each year"

fig_how_many_jobs_per_year <- 
ggplot(jobdata) +
  aes(year_ad_posted) +
  geom_bar() +
  scale_x_discrete(name = "") +
  ylab("Number of\ntenure track job ads") +
  theme_minimal(base_size = 28) +
  guides(x = "none")

@fig-how-many-jobs-per-year shows how many jobs per year in our sample

```{r}
#| label: fig-how-many-jobs-per-year-by-rank
#| fig-cap: "Proportion of tenure-track archaeology faculty jobs per year by rank"

# explore over time
fig_prop_by_job_title_per_year <- 
jobdata %>%
  group_by(year_ad_posted) %>%
  count(job_title_simple) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot() +
  aes(year_ad_posted,
      prop,
      group = job_title_simple,
      colour = job_title_simple) +
  geom_line(linewidth = 2) +
  theme_minimal(base_size = 28) +
  xlab("") +
  ylab("Proportion of\nall tenure-track ads") +
  theme(legend.position = c(0.5, 0.5)) +
  scale_colour_brewer(palette = "Dark2") +
  guides(colour = guide_legend(nrow=2,
                               byrow=TRUE,
                               "Job title")) +
  guides(x = "none")

# ratio of tenure-track to untenured positions
# base URL changes after 2018_2019

# make a set of URLs for each page for each year
urls_for_each_year <- c(str_glue('{base_url_to_2019}{years_to_2019}'), 
                        str_glue('{base_url_after_2020}{years_after_2020}')) %>% 
    str_replace("2021-2022", "2021-22")

# count of TT jobs per year from our manual data collection,
# join with our total number of all jobs by scraping
count_of_tt_jobs_per_year_from_our_form <- 
jobdata %>% 
  group_by(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page) %>% 
  tally() %>% 
  right_join(total_number_of_jobs_per_year_tbl) %>% 
  rename(n_tt_jobs  = n,
         n_total_jobs = total_number_of_jobs_per_year) %>% 
  mutate(n_non_tt_jobs = n_total_jobs - n_tt_jobs,
         ratio_tt_2_ntt = n_tt_jobs / n_non_tt_jobs) %>% 
  mutate(year = str_extract(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page,  "[[0-9]]{4}-[[0-9]]{4}|2021-22")) %>%  
  mutate(year = ifelse(year =="2021-22",  "2021-2022", year))

# draw plot
fig_ratio_tt_2_ntt_jobs_per_year <- 
  ggplot(count_of_tt_jobs_per_year_from_our_form) +
  aes(year, 
      group = 1,
      ratio_tt_2_ntt) +
  geom_line(linewidth = 2) +
  geom_hline(yintercept = 1,
             colour = "red") +
  annotate("text", 
           x = 3, 
           y = 1.3, 
           label = "1:1 ratio",
           colour = "red") +
  labs(y = "Ratio of tenure-track\nto non-tenure track and other",
       x = "") +
  theme_minimal(base_size = 28) +
  scale_x_discrete(name = "Academic Year") 
   #guides(x = "none")

```{r}
# save these three plots as one set
library(cowplot)
plot_grid(
  fig_how_many_jobs_per_year,
  fig_prop_by_job_title_per_year,
  fig_ratio_tt_2_ntt_jobs_per_year,

  ncol = 1,
  align = "hv",
  axis = "lr",
  labels = "AUTO"
)

>>>>

title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 900
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Academic careers are frequently sought by archaeology graduate students. Job listing websites often serve as the first place for these students when seeking academic positions. We examined tenure-track job advertisements over the past decade to gain insights into the academic job market for archaeologists. Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time. We studied the text of 449 job ads posted from 2013-2023. Our analysis focuses on shifts in archaeological topics and methods requested in job ads. We investigate whether the burden on applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago? We also examine whether there is an increasing trend in job advertisements highlighting diversity and inclusivity, thereby encouraging a broader range of applicants. Additionally, we assess the influence of socio-political factors on the changing focus of research topics in the field. This research aims to assist current and future archaeology students and graduates in better understanding the job market and the requirements of employers, thereby aiding them in effectively preparing for their applications for positions in archaeology.
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States. We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.

# Background

@@

total_number_of_ads_in_our_sample <- nrow(jobdata) # 550

# for text on our figures
base_size <- 6
```

```{r}
#| label: fig-how-many-jobs-per-year

```{r}
#| label: fig-how-many-jobs-per-year-by-rank

# explore over time
fig_prop_by_job_title_per_year <- 
jobdata %>%
  group_by(year_ad_posted) %>%
  count(job_title_simple) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot() +
  aes(year_ad_posted,
      n,
      group = job_title_simple,
      fill = job_title_simple) +
 # geom_line(linewidth = 2) +
  geom_col() +
 theme_minimal(base_size = base_size) +
  xlab("") +
  ylab("Number of\ntenure track job ads") +
  scale_fill_brewer(palette = "Dark2") +
  guides(fill = guide_legend(nrow=2,
                               byrow=TRUE,
                               "Job title",
                                keywidth=0.1,
                 keyheight=0.1,
                 default.unit="inch")) +
    theme(legend.position = c(0.5, 0.85)) +
  guides(x = "none")

# make a set of URLs for each page for each year
urls_for_each_year <- c(str_glue('{base_url_to_2019}{years_to_2019}'), 
                        str_glue('{base_url_after_2020}{years_after_2020}')) %>% 
    str_replace("2021-2022", "2021-22")
```

```{r}
#| eval: false

# scrape the web pages to get the 
# ratio of tenure-track to untenured positions
# base URL changes after 2018_2019
# We did this once and then saved the output so we don't scrape every
# time the quarto document is rendered. So this code block is
# eval: false to avoid repeated and unnesecary scraping

# save this list so we don't have to scrape again
library(rlist)
list.save(urls_for_each_year_headers, 
          here('analysis/data/raw_data/urls_for_each_year_headers.yaml'))
```

```{r}
#| label: ratio-tt-to-non-tt-jobs

library(rlist)
urls_for_each_year_headers <-  list.load( here('analysis/data/raw_data/urls_for_each_year_headers.yaml'))

# count of TT jobs per year from our manual data collection,
# join with our total number of all jobs by scraping
count_of_tt_jobs_per_year_from_our_form <- 
jobdata %>% 
  group_by(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page) %>% 
  tally() %>% 
  right_join(total_number_of_jobs_per_year_tbl) %>% 
  rename(n_tt_jobs  = n,
         n_total_jobs = total_number_of_jobs_per_year) %>% 
  mutate(n_non_tt_jobs = n_total_jobs - n_tt_jobs,
         ratio_tt_2_ntt = n_tt_jobs / n_non_tt_jobs) %>% 
  mutate(year = str_extract(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page,  "[[0-9]]{4}-[[0-9]]{4}|2021-22")) %>%  
  mutate(year = ifelse(year =="2021-22",  "2021-2022", year))  %>% 
  # add line break for prettier plotting
  mutate(year_ad_posted_break = str_replace(year, "-", "-\n"))

# draw plot
fig_ratio_tt_2_ntt_jobs_per_year <- 
  ggplot(count_of_tt_jobs_per_year_from_our_form) +
  aes(year_ad_posted_break, 
      group = 1,
      ratio_tt_2_ntt) +
  #geom_line(linewidth = 2) +
  geom_col() +
  geom_hline(yintercept = 1,
             colour = "red") +
  annotate("text", 
           x = 6, 
           y = 1.4, 
           label = "1:1 ratio",
           colour = "red",
           size = 3) +
  labs(y = "Ratio of tenure-track\nto non-tenure track and other",
       x = "") +
  theme_minimal(base_size = base_size) +
  scale_x_discrete(name = "Academic Year") 
   #guides(x = "none")

```{r}
#| label: combine-three-basic-plots

# save these three plots as one set
library(cowplot)
plot_grid(
  fig_prop_by_job_title_per_year,
  fig_ratio_tt_2_ntt_jobs_per_year,

  ncol = 1,
  align = "hv",
  axis = "lr",
  labels = "AUTO",
  label_size = 8
)

```{r}
#| label: show-basic-plots
#| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank. B: Ratio of tenure-track to non-tenure-track positions over time."

# show figure from PNG file
knitr::include_graphics("../figures/fig-panel-per-year.png")
```

We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023. Figure @show-basic-plots shows the count of ads for each year. Assistant Professor jobs are consistently the most common title and rank, and open rank or full professor the least frequent. The ratio of tenure-track to non-tenure track positions is generally well above one, with

@@

```{r}
# stop rendering at this point because the code below is not working
knitr::knit_exit()
```
